A Component-Based Sensor Network for \\ Environmental Monitoring

### 2009-03-06 docs/icccn2009/paper.tex
More text
git-svn-id: svn://localhost/netbeams/trunk@349 559edde2-a4d4-44e0-96fc-4035e40d7396

\title{An OSGi-based Sensor Network for \\ Environmental Monitoring}

@@

\cite{roemer:2004}

@@

\section{Background}

@@

The YSI 6600EDS V2 Sonde is a water quality monitoring device that
gathers water quality data, and operates in fresh, sea, or
polluted water. It is under 22 inches (under 55 cm) in length, 3.5
inches (8.9 cm) in diameter, and weighs approximately 7 pounds (3.18
kg). It operates at temperatures between -5 to +50 deg C and at depths
up to 656 feet (200 meters).  It uses 8 C-size Alkaline Batteries or
External 12 VDC. Battery life can last up to 75 days depending on
sensor configuration. The Sonde's bulkhead contains multiple pin
connectors to support sensor probes for measuring the following:

@@

Dissolved oxygen
Conductivity/Temperature
Fouling Resistant pH/ORP
Chloride
Ammonium
Nitrate
Turbidity
Chlorophyll
Rhodamine
PC-Blue-green Algae
PE-Blue-gren Algae

@@

Start with some related work. Who else has done an OSGi-based
infrastructure for sensor networks? Kleber: I believe you have a few
references. This section should be a combination of a use case (using
the RTC and their YSI sonde as an example) and a requirements
analysis: requirements such as remote management and monitoring
capabilities, time-delayed communication, etc. In this section should
be no mentioned of DSP or OSGi. Length: 1 page.

@@

Should describe the OSGi-based NetBEAMS architecture. Total length of
this section: 2 pages.

Micro-kernel architecture.

@@

In order to enrich the DSP, a programmer has to implement new DSP
Components.

Should give some API sniplets for DP and DC.

@@

@@

@@

Should explain the way DSP Messages are defined: XML Schema and JAXB.

@@

@@

In the following we present a practical application the the DSP
presented in the previous section. We describe a deployment of the DSP
using sensor owned and operated by the RTC. Our environmental sensor
network is dubbed NetBEAMS (Networked Bay Environmental Assessment
Monitoring System) and allows access to sensor equipment deployed
throughout the San Francisco Bay Area. Figure \ref{FIG_NETBEAMS}
provides an end-to-end overview of the NetBEAMS architecture. NetBEAMS
is built using multiple installtion of the DSP. One DSP installation
is collocated with each sensor and an additional DSP is running on the
backend server.

@@

To describe NetBEAMS is to describe the specific DC and DP we have
implemented. On the far left of Figure \ref{FIG_NETBEAMS} is the
actual sensor; in our case the aforementioned Seabird sensor.
Collocated with the Seabird is an embedded system. Here we make use of
a platform called Gumstix.

The Gumstix is a single-board computers with 256 Mb of flash and 256
Mb of SDRAM, and running a 600 MHz TI OMAP 3503 processor. Additional
features can be added to the motherboard with expansion cards
connected via on-board buses. The motherboards draw less than 250 mA
@4V at 400 MHz and less than 50 mA while idling, waiting for input.
Gumstix run Linux 2.6 with the BusyBox utilities, and use the
OpenEmbedded build environment to provide a complete Linux
environment and a large range of Linux applications. We use the JamVM
implementation of Sun Microsystem's virtual machine to execute Java
programs.

The Huawei E220 is a USB modem that supports HSDPA, UMTS, EDGE packet
data services at maximum transmission rates of 3.6Mbps, 384kbps, and
236.8kbps respectively. HSDPA and UMTS operate at 2100MHz while GSM,
GPRS, and EDGE operate at 900, 1800, and 1900MHz. The modem connects
to the Verdex Pro XM4 motherboard via a mini USB interface (supporting
USB 2.0). Linux kernels since 2.6.20 include support for E220 drivers.
This includes the linux kernel on the gumstix which runs 2.6.21.  The
gumstix communicates with the modem using the PPP protocol and
accesses the cellular network using a data plan from AT\&T.

\subsection{YSI Sonde DP}

@@

\subsection{Wire transport DC/DP}

\subsection{Web Managements}

@@

Length (including bibliography): 0.5 pages.

>>>>

\title{A Component-Based Sensor Network for \\ Environmental Monitoring}

@@

Mention the nature of our sensor network: highly specialized sensors,
sparse topology, remote locations, long range communication.
\cite{roemer:2004}

@@

\section{Background}
\label{SEC_BACKGROUND}

@@

The YSI 6600EDS V2 Sonde is a water quality monitoring device that
gathers water quality data, and operates in fresh, sea, or polluted
water. Some of the measurements that the YSI is capable of are
conductivity, temperature, chloride, ammonium, nitrate, turbidity, and
chlorophyll.  It is under 55 cm in length, 8.9 cm in diameter, and
weighs approximately 3.18 kg. It operates at temperatures between -5
to +50 deg C and at depths up to 200 meters.  It uses 8 C-size
Alkaline Batteries or External 12 VDC. Battery life can last up to 75
days depending on sensor configuration. The Sonde's bulkhead contains
multiple pin connectors to support sensor probes for measuring the
following:

@@

@@

Start with some related work. Who else has done an OSGi-based
infrastructure for sensor networks? Kleber: I believe you have a few
references. This section should be a combination of a use case (using
the RTC and their YSI sonde as an example) and a requirements
analysis: requirements such as remote management and monitoring
capabilities, time-delayed communication, etc. In this section should
be no mentioning of DSP. Length: 1 page.

@@

Should describe DSP architecture. Only mention OSGi in last subsection
and explain that components can be mapped to bundles. Currently no
distinction between Platform and Framework. Necessary?

@@

All DSP Components need to implement the following Java interface:

@@

By implementing the above Java interface, the resuling code
effectively becomes a DSP Component that can be deployed at runtime. A
component is identified by its type. Only one component of a certain
type per DSP is permissible. A string serves as a type identifier that
much be unique across all DSP Components. Upon deploying a DSP
Component, an initialization function is called (line 4). Once a
component is initialized, it may be started (line 5) and stopped (line
6) multiple time (TODO: SHOULDN'T THERE BE ALSO A WAY TO UNDEPLOY A
COMPONENT?) Whenever the DSP wishes to pass a message (that was
generated by some DP), it invokes the \texttt{deliver()} method (ine
7). The message structure will be explained in the following section.

@@

Message passing is the main paradigm for exchanging information within
a DSP. A data producer sends a message with data consumers accept
messages. We make use of XML technologies to define the format of a
message. Specifically, the common header of all messages are defined
via the following XML Schema:

@@

The above XML Schema DSPMessage serves as a base schema for all
messages and itself defines information typically contained in the
header of a protocol data unit. Amongst others, the header contains a
unique message ID (line 8) and the originating DP (line 9). Using the
XML Schema, we use JAXB to generate a type safe API for handling
instances of DSPMessage. Different messages with different payloads
can be defined by creating schemas that derive from DSPMessage. While
inside the DSP all messages are of type DSPMessage, specific DP and DC
can distinguish between different message types based on their runtime
type information.

@@

can only route messages within one DSP. No notion of remote DSP
components.

\subsection{OSGi}

OSGi lends itself naturally to implement the DSP.

DSP can be implemented in different ways, but OSGi is a natural match.
DSP Components are mapped to bundles. DSP itself is a bundle.

@@

In the following we present a practical application of the DSP
presented in the previous section. We describe a deployment of the DSP
using sensor owned and operated by the RTC. Our environmental sensor
network is dubbed NetBEAMS (Networked Bay Environmental Assessment
Monitoring System) and allows access to sensor equipment deployed
throughout the San Francisco Bay Area.  Our goal is to address the
problem of data collection from remote sensors by using the DSP to
automatically interrogate sensors at the remote field locations and
transmit the collected measurements over long distances to a backend
server where the data is made available to potential users in near
real-time. This end-to-end system will be a ready-to-use turnkey
solution that is adaptable to a wide variety of sensor types and
requires no programming expertise to operate. Configuration of the
system will be possible when the sensor is deployed in the field via a
web-based interface, eliminating the need for extra site visits by
technicians between service cycles to change systems functionalities
(e.g. sampling rate). We present the NetBEAMS infrastructure by
discussing its hardware and software components.

\subsection{Hardware}

Figure \ref{FIG_NETBEAMS} provides an end-to-end overview of the
NetBEAMS architecture. NetBEAMS is built using multiple installation
of the DSP and by providing special purpose components for the various
tasks. One DSP installation is collocated with each sensor and an
additional DSP is running on the backend server. Because of the remote
locations where each individual sensor is deployed in the field and
the sparse nature of the network, the resulting topology is a tree
structure of depth 1.

@@

NetBEAMS supports environmental sensor such as the YSI Sonde
introduced in Section \ref{SEC_BACKGROUND}. In order to access its
sensor readings, we collocate it with an off-the-shelf embedded
computing platform called Gumstix (REF) (far left in Figure
\ref{FIG_NETBEAMS}). The Gumstix is a single-board computers with 256
Mb of flash and 256 Mb of SDRAM, and running a 600 MHz TI OMAP 3503
processor. Additional features can be added to the motherboard with
expansion cards connected via on-board buses. The motherboards draw
less than 250 mA @4V at 400 MHz and less than 50 mA while idling,
waiting for input.  Gumstix runs Linux 2.6 with the BusyBox utilities,
and use the OpenEmbedded build environment to provide a complete Linux
environment and a large range of Linux applications. We use the JamVM
(REF) implementation of Sun Microsystem's virtual machine in order to
run the DSP

We make use of the cellular phone network to establish a communication
link with the backend server. The cellular phone network has the
advantages of reaching well beyond telemetry options such as wireless
networks and being a significantly lower cost solution than satellite
communications. We make use of the Huawei E220. The E220 is a USB
modem that supports HSDPA, UMTS, EDGE packet data services at maximum
transmission rates of 3.6Mbps, 384kbps, and 236.8kbps respectively.
HSDPA and UMTS operate at 2100MHz while GSM, GPRS, and EDGE operate at
900, 1800, and 1900MHz. The modem connects to the Verdex Pro XM4
motherboard via a mini USB interface (supporting USB 2.0). Linux
kernels since 2.6.20 include support for E220 drivers.  This includes
the linux kernel on the Gumstix which runs 2.6.21.  The Gumstix
communicates with the modem using the PPP protocol and accesses the
cellular network using a data plan from AT\&T.

\subsection{YSI Sonde Component}

Our implementation of the DSP is based on the Open Source
implementation of OSGi called Knopflerfish (REF). Since Sun
Microsystems does not support its JDK for the Gumstix, we use JamVM
(REF) as the Java virtual machine implementation and GNU Classpath
(REF) for the runtime libraries. We have implemented various DSP
Components that implement the functionality needed for NetBEAMS. On
the far left of Figure \ref{FIG_NETBEAMS} is the actual sensor; in our
case the aforementioned YSI Sonde. A YSI Sonde data procuder (1) is a DSP
component that uses the RS232 to read the physical measurements from
the device. This component converts the physical measurement into a
DSP-specific message. The following XML Schema defines the YSI Sonde
specific message that captures all measurement points mentioned in
Table \ref{TAB_SONDE_MEASUREMENTS}:

@@

The schema is derived from DSPMessage. As a payload it adds an
unbounded sequence of sondeDataType. Each instance of this type
captures one timestamped measure point. The YSI Sonde component can be
configured via special DSP messages that define properties such as the
sampling frequency. These property messages originate from the web
frontend that we will explain in a subsequent section.

\subsection{Remote Communication}

The YSI Sonde component will pass any messages it generates to the
local DSP. Via the matcher configuration these messages are forwarded
to the remote communication component (2). It is the responsibility of
this component to upload the message to its remote peer component
(3). Since we use a data channel of a cellular phone network, the
remote communication component only uploads the buffered data in
certain configurable time intervals.

\subsection{Web Management}

Web Interface. Management messages are regular
DSPMessages. Asynchronous UI: UI may only build up after next
communication cycle.

@@
